Formal Verification of Forward Secrecy and Post-Compromise Security for TreeKEM

### 2024-10-05 manuscript/manuscript-SBMF.tex
Addressing SBMF reviewer comments

\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The TreeKEM protocol is the preeminent implementation candidate for the Message Layer Security standard.
Prior work analyzed TreeKEM by defining the Continuous Group Key Agreement security game,
which facilitated proof of some security guarantees and also identified protocol deficiencies which were subsequently remedied.
This work extends such applications by formalizing the Continuous Group Key Agreement security game through multiple soundness preserving abstractions.
The model is parameterized by $N$, representing an unbounded protocol duration between at most $N$ distinct participants.
Once formalized, the game is encoded within \Promela and essential security guarantees are verified for $N \le 16$ via the model checker Spin.
This represents a notable achievement, both in practical security terms for the TreeKEM protocol, as well as demonstrating scalability techniques for non-trivial parameter bounds when modeling a complex, concurrent protocol.
\keywords{Cryptographic Protocols, Formal Verification, Linear Temporal Logic, Model Checking, \Promela, Spin}
\end{abstract}

@@

The TreeKEM protocol \cite{bhargavan:hal-02425247} was conceived in 2018, and thoroughly described in 2019.
%[SSB] Was the description English? formal? If informal, it might be a good place to bring the games in (not sure). 
The \Abrev{IETF} along with many other corporate and government sponsors have put their support behind the TreeKEM protocol.
TreeKEM provides functionality for achieving each of the six operations to be considered a \Abrev{SGM} protocol.
%[SSB] these ops are a forward ref. Start by saying it requires 6 ops, and perhaps a parenthethical remark that these are discussed in Sec foo. 
Additionally TreeKEM aims to satisfy both the efficiency goals and security guarantees of the \Abrev{MLS} specification, addressing the previously open problems of \Abrev{FS} and \Abrev{PCS} for \Abrev{SGM}.
Conformance with both of these definitions is a result of TreeKEM's construction.

@@

%[SSB] This seems to be defining two different terms (epoch and communication epoch) with two definitions. Use one term, one def. 
The TreeKEM protocol is temporally delineated into discrete quanta referred to as ``epochs,'' representing a collection of messages sent using the same secret root key.
% [SS] "evolve their" sounds weird for a shared subtree. Should it be "evolve the [shared]"?
% [AW] sure thing!
To create a new communication epoch, any group member may broadcast a control message for distribution via the \Abrev{DS}, instructing the other group members how to evolve the shared \Abrev{LBBT} and associated key material to arrive at the shared symmetric key of the new epoch.
Figure \ref{fig:SGM-Timeline} depicts a simple case of communication epoch progression.
New encryption ``epochs'' are created by a member broadcasting a control message for one of the following operations:

@@

Prior work has analyzed TreeKEM through a security game~\cite{alwen2020security}, which will be further elaborated on and utilized below in Section \ref{sec:Formal-Abstractions}.
The resulting analysis yielded a mathematical proof, via graph theory, of the security guarantees \Abrev{FSU} and  \Abrev{PCS} for TreeKEM.
In contrast our work presents a mechanical verification of the same security guarantees for TreeKEM, up to some parameterized bound,  by utilizing model checking.

@@

Formalizing the model under verification is always an important and detailed process.
The verification methodology chosen by the authors utilizes explicit state model checking.
For this reason, it is important to rigorously control the size of the model's state-space in order to avoid a state-space explosion. 
To that end, the authors developed multiple abstractions (presented in Section~\ref{sec:Formal-Abstractions}) which, when composed together progressively reduce the model's state-space. 
Without these abstractions, the model is not verifiable for even modestly sized parameters.

\subsection{Continuous Group Key Agreement}
In the two-party case, a general notion of Continuous Key Agreement has been used to provide robust security guarantees such as forward secrecy and post-compromise security~\cite{perrin2013axolotl,perrin2016double,alwen2019double}.
For \Abrev{SGM}, this same notion has been extended as \Abrev{CGKA}, which is used as an abstraction for reasoning about the \Abrev{MLS} security guarantees~\cite{alwen2020security}.
A continuous group key-agreement scheme \CGKAdef\ is an abstract collection of algorithms used to maintain secure communications between a group of two or more participants, each algorithm being known and performed locally by participants:\\[2mm]
\centerline{$\CGKAdef = \left\{\,\Protocol{init},\, \Protocol{create},\, \Protocol{add},\, \Protocol{rem},\, \Protocol{upd},\, \Protocol{proc}\,\right\}$}\\[-3mm]

@@

These oracles define a security game \cite{shoup2004sequences} which emulates the \Abrev{ITM}.
Such security games are a common analysis tool made by cryptographers, where the ``game moves'' representing protocol semantics, and the objective is to show that the probability of an adversary \Adversary  breaking a given security property by playing this game is negligible.
Within the \CGKAsec, an adversary \Adversary is parameterized by \((T, C, N)\).
The \(T\) value indicates that the protocol runs in at most \(T\) epochs, during which time \(\mathcal{A}\) can make at most \(C\) ``challenge queries'' to the \Oracle{reveal}{} or \Oracle{chall}{} oracles, and the total unique group members of the protocol can never exceed \(N\).
The goal of \Adversary is to infer information about the shared TreeKEM root key.
Within the security game, this is modeled by \Adversary  ending the game via a challenge query to the \Oracle{chall}{} oracle, 
and subsequently demonstrating non-negligible advantage.
%[SSB] Define advantage ('(\ie ...)')
We model the TreeKEM protocol in terms of the \CGKAsec as the initial abstraction.
%[SSB] Its not clear that following subsections are abstractions. Either/both add a sentence here saying they are, or change subsection names to "Abstraction: ...". Maybe also mention here that 'this model needs blah hours to verify even for N=4' if you have such numbers.

@@

\subsection{Adversarial Direction}
While both the adversary \Adversary and each group member can initiate a new epoch within the \CGKAsec, modeling this explicitly unnecessarily complicates the model.
Recall that within the \Abrev{ITM}, \Adversary controls the network.
This is modeled within the \CGKAsec by messages sent from any source only being delivered if \Adversary permits it by querying \Oracle{deliver}{}.
Because the verification methodology will explore all possible interleaving of concurrent actions within the \CGKAsec, simulating nondeterministic computation, this can be utilized to form an additional simplifying abstraction.

Suppose for a given epoch $e_i$ that the non-empty set $M$ consists of all ``epoch advancing messages'' originating from calls to algorithms \Protocol{add}, \Protocol{rem}, and \Protocol{upd} by group members.
By the rules of the \CGKAsec, \Adversary selects some message $m \in M$ sent by group member $j$ and queries \Oracle{deliver}{g_j,\,g_k} to propagate delivery of $m$ to the group members and advance to the next communication epoch.
Without loss of generality, rather than group member $j$ creating and sending $m$ by calling either algorithm \Protocol{add}, \Protocol{rem}, or \Protocol{upd}; the adversary \Adversary equivalently queries the corresponding \Oracle{add-user}{g_j,\,g_k}, \Oracle{remove-user}{g_j,\,g_k}, or \Oracle{send-update}{g_j,\,g_k} oracle, respectively.
This operational equivalence permits the modeling of the \CGKAsec to only consider \Adversary--originating calls to algorithms \Protocol{add}, \Protocol{rem}, and \Protocol{upd} via queries to the oracles \Oracle{add-user}{}, \Oracle{remove-user}{}, and \Oracle{send-update}{}, respectively.
Consequently, the model does not need to consider any group member calls to \Protocol{add}, \Protocol{rem}, and \Protocol{upd}, allowing these states to be safely removed from the model encoding without effecting protocol semantics.
The model can explicitly track epoch progress by the adversary via $t$ and $c$ variables corresponding to epoch index and a binary indicator of a challenge query made within the current epoch, respectively.
%The application of this simplifying abstraction composed with the original \CGKAsec is depicted in Figure \ref{fig:CGKA-Adversarial-Driven}.
% [SS] I think the main point of above paragraph is that modeling/proving by epoch avoids the need for non-A originating calls? But I'm not sure its clear to reader. 
%[AW] Some more descriptive language here.
%[SSB] This sounds like partial order reduction (enabled by default in SPIN), but can be disabled using -DNOREDUCE. Presumably our solution improved over SPIN - assuming so add sentence here saying so. Also check SPIN docs to ensure that its actually using the reduction (it doesnt in some cases).

@@

\subsection{Parameterized Bounds of $T$, $C$, and $N$}

@@

\subsection{Required Progression\label{sec:abstraction-progression}}

@@

\subsection{Markovian Property}

@@

The \Promela encoding \footnote{Source code: \url{https://github.com/recursion-ninja/Manuscript-2024-SBMF/model}} of \CGKAmodN{N} produces all possible states and transitions of TreeKEM, and this representation is then supplied to a model checker for verification.

The explicit state model checker Spin~\cite{HolzmannSpinBook2003} was chosen as the verification tool for this work.
As an explicit-state model checker, Spin utilizes a representation of all possible model states as a \Abrev{FSM}~\cite{clarke1981design}, with transitions corresponding to possible changes in model state.
Spin then asserts that a property holds for the protocol by traversing through all combinations of transitions through the graph.
%~\cite{Vardi1986}.
The principle drawback of explicit state model checking is the immense state-space resulting as the model complexity grows~\cite{burch1992symbolic}.
As a venerable tool, Spin does much to address this~\cite{rudin1987limits} but the authors' own encoding adaptations which further mitigate the state-space explosion are detailed below in Sections \ref{sec:bit-array} and \ref{sec:popcount}.
Additionally, Spin supports numerous compile time directive which alter the state-space representation, search trajectories, and time-memory trade-offs (see Table \ref{tab:spin-confg}).

@@

In the default \Promela type system, an array of bits is encoded as an array of bytes at runtime, resulting in each ``bit value'' requiring 8 bits during verification.
This encoding inflates the state-vector size unnecessarily by 8 bits and causes the model checker to consider 256 states instead of the actual 2 binary states which are valid encodings ($\texttt{0x00} = \textsc{True}$, $\texttt{0xFF} = \textsc{False}$).
Consequently, verification times are slower, and at least eight times more memory is required.
To address this, 
bits within a fixed bit-width \Promela type are manually toggled using C Pre-processor macros,
% [SS] Check that my rephrasing is correct
% [AW] Yes!
%the manual toggling bits within a fixed bit-width \Promela type using C Pre-processor (CPP) macros is included, 
reducing the state-vector size and improving runtime efficiency while minimizing memory usage.

@@

The authors included PopCount as a small, inlined sequence of arithmetic and bit-wise operations which runs in \BigO{1} time with no additional memory.
Substituting PopCount in place of the naive iteration has two notable effects, reducing the state-vector length by removing the loop counter and reducing the number of state transitions by removing the entire control flow structure of the loop.
For a bit-array of length $n$, the number of transitions encoded in the model by iteration is exponentially related to $n$, whereas the PopCount macro derives a \emph{constant} number of transitions independent of $n$.

@@

%[SSB] what does UnderNoteRef mean? If there is a point to it, define what it means; else get rid of it. 
\begin{definition}[Forward Secrecy with Updates]
If the state of any group member is leaked at some during an epoch\UnderNoteRef{1}, the shared keys from all previous epochs remain hidden from the adversary\UnderNoteRef{2}.
\end{definition}%
\noindent\textbf{LTL Predicate} \LTLPredicate{FSU}\textbf{:}%
{\small%
\[%
\Box \big(\, \underset{\UnderNote{1}}{\texttt{CGKA@move\_corrupt}} \implies \underset{\UnderNote{2}}{\texttt{learnedLegacyKey}} = \bot \,\big)%
\]%
}

The \Promela model is named \TermLTL{CGKA}, and the state within the model at which \Adversary querries the oracle  \Oracle{corr}{\TermLTL{targetID}} is labeled \TermLTL{move\_corrupt}.
The term \TermLTL{CGKA@move\_corrupt} will be true only when \TermLTL{CGKA} is in state \TermLTL{move\_corrupt}.
Additionally, the model contains a boolean flag \TermLTL{learnedLegacyKey} indicating when \Adversary has learned, during the \emph{current} epoch, information about the shared group key of any \emph{any previous} epoch.
The \Abrev{LTL} formula \LTLPredicate{FSU} is a direct translation of the natural language \Abrev{FSU} description.

@@

The methodology presented is two-fold.
First, the performance is measured of the model \CGKAmodN{N} described in this work and compared with the performance of a precursor model defined by~\cite{washburn2022formal} 
denoted as \CGKAmodNT{N}{T}.
%[SSB] after "described" add "in sections 2.2-2.6"
The comparison model \CGKAmodNT{N}{T} is the same semantic encoding of \CGKAmodN{N} \emph{before} the abstractions described were applied, i.e:  $$\Code{Abstractions}\Parens{\CGKAmodNT{N}{T}} =  \CGKAmodN{N}$$
Therefore \CGKAmodNT{N}{T} is an apt choice as a comparative baseline from which the effects of the composed abstractions can be measured.
Comparison of performance is made across several metrics, including state-vector length, runtime, memory usage, state-space size, and transition count.
Both models performed verification of the \LTLPredicate{FSU} and \LTLPredicate{PCS} properties.
Second, \CGKAmodN{N} is utilized to verify \LTLPredicate{FSU} and \LTLPredicate{PCS} for parameters far beyond the the capabilities of \CGKAmodNT{N}{T}.
Subsequently, the authors provide brief contextual commentary accompanying these verification results.

@@

\centering

@@

\hfil

@@

& \multicolumn{2}{c}{\TableHeadFontSize\parbox[][6mm][c]{13mm}{\centering \CGKAmodN{N}}}
& \multicolumn{2}{c}{\TableHeadFontSize\parbox[][6mm][c]{15mm}{\centering \CGKAmodNT{N}{T}}}\\

@@

& \parbox[t]{4mm}{\raggedleft$T\:$}
& \parbox[t]{8mm}{\scriptsize \centering State\\Vector}
& \parbox[t]{4mm}{\raggedleft$T$}
& \parbox[t]{8mm}{\scriptsize\centering State\\Vector}\\

@@

\hfil

@@

& \parbox[t]{7mm}{\TableHeadFontSize\centering $N$}

@@

Given the new range of tractable verification provided by \CGKAmodN{N}, verification was performed with model parameters which were previously untenable using \CGKAmodNT{N}{T}. 
Verification of  \Abrev{FSU} and \Abrev{PCS} was conducted on the models \CGKAmodN{16}.
Both models had a state-vector of $100$\siBytes.
Then verification results produced by  \CGKAmodN{16} for both properties were in the \emph{affirmative}.
The positive verification of the LTL encodings for \Abrev{FSU} and \Abrev{PCS} given by \CGKAmodN{16} are not surprising.
Rather, it is consistent with the expectations shown by the previous \CGKAdef abstraction formalization work~\cite{alwen2020security}.
While consistent, this is the first instance known to the authors which has modeled the TreeKEM protocol and verified \Abrev{FSU} and \Abrev{PCS} for all $(\,T, C, N\,)$--Adversaries, with $T \in \NumericRangeOpenR{1}{\infty}$, $C \in \NumericRange{1}{T}$, and $N \in \NumericRange{2}{16}$.
Formal verification for TreeKEM of this scope has not been done before.

@@

This work presents a verification of \Abrev{FSU} and \Abrev{PCS} for TreeKEM; an important undertaking in establishing the security of and confidence in the the protocol.
Furthermore, an example of effective, and generally applicable model abstraction and model encoding techniques is illustrated.
The authors believe that, though no abstraction in isolation is novel, this is the first work to compose these abstractions and encoding techniques in the combination presented.
The exploration of combining these techniques has been shown necessary for achieving scalability when verifying a complex model such as the TreeKEM protocol.
An unqualified improvement was obtained across each of the following measurable dimensions: model complexity,  parameterization, state-vector length, and verification space as well as time.
These improvements were essential to perform verification of the largest known TreeKEM group size to date.
While the application and analysis of presented techniques herein was limited to TreeKEM, the techniques are not so bespoke that they apply only to modeling and verifying TreeKEM.
Rather, it is probable that only minor creativity is require to transfer them to other modeling efforts.

>>>>

\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The TreeKEM protocol is the preeminent implementation candidate for the Message Layer Security standard.
Prior work analyzed TreeKEM by defining the Continuous Group Key Agreement security game,
which facilitated proof of some security guarantees and also identified protocol deficiencies which were subsequently remedied.
This work extends such applications by formalizing the Continuous Group Key Agreement security game through multiple soundness preserving abstractions.
The model is parameterized by $N$, representing an unbounded protocol duration among $N$ distinct participants.
Once formalized, the game is encoded within \Promela and essential security guarantees are verified for $N \le 16$ via the model checker Spin.
This represents a notable achievement, both in practical security terms for the TreeKEM protocol, as well as demonstrating scalability techniques for non-trivial parameter bounds when modeling a complex, concurrent protocol.
\keywords{Cryptographic Protocols, Formal Verification, Linear Temporal Logic, Model Checking, \Promela, Spin}
\end{abstract}

@@

The TreeKEM protocol \cite{bhargavan:hal-02425247} was (informally) conceived in 2018, and formally described in 2019.
%[SSB] Was the description English? formal? If informal, it might be a good place to bring the games in (not sure). 
The \Abrev{IETF} along with many other corporate and government sponsors have put their support behind the TreeKEM protocol.
TreeKEM provides functionality for achieving each of the six operations to be considered a \Abrev{SGM} protocol (discussed further in Section \ref{sec:CGKA}).
%[SSB] these ops are a forward ref. Start by saying it requires 6 ops, and perhaps a parenthethical remark that these are discussed in Sec foo. 
Additionally TreeKEM aims to satisfy both the efficiency goals and security guarantees of the \Abrev{MLS} specification, addressing the previously open problems of \Abrev{FS} and \Abrev{PCS} for \Abrev{SGM}.
Conformance with both of these definitions is a result of TreeKEM's construction.

@@

%[SSB] This seems to be defining two different terms (epoch and communication epoch) with two definitions. Use one term, one def. 
Correspondingly, the TreeKEM protocol is temporally delineated into discrete quanta referred to as ``communication epochs,'' representing a collection of messages sent using the same secret root key.
% [SS] "evolve their" sounds weird for a shared subtree. Should it be "evolve the [shared]"?
% [AW] sure thing!
To create a new communication epoch, any group member may broadcast a control message for distribution via the \Abrev{DS}, instructing the other group members how to evolve the shared \Abrev{LBBT} and associated key material to arrive at the shared symmetric key of the new epoch.
Figure \ref{fig:SGM-Timeline} depicts a simple case of communication epoch progression.
New encryption ``epochs'' are created by a member broadcasting a control message for one of the following operations:

@@

Prior work has analyzed TreeKEM through a security game~\cite{alwen2020security}, which will be further elaborated on and utilized below in Section \ref{sec:Formal-Abstractions}.
The resulting analysis yielded a mathematical proof, via graph theory, of the security guarantees \Abrev{FSU} and  \Abrev{PCS} for TreeKEM.
In contrast, presents a mechanical verification of the same security guarantees for TreeKEM, up to some parameterized bound,  by utilizing model checking.

@@

Formalizing the model under verification is always an important and detailed process.
The verification methodology chosen by the authors utilizes explicit state model checking.
For this reason, it is important to rigorously control the size of the model's state-space in order to avoid a state-space explosion. 
To that end, the authors developed multiple abstractions (presented in Section~\ref{sec:Formal-Abstractions}) together, progressively reduce the model's state-space. 
Without these abstractions, the model is not verifiable for even modestly sized parameters.

\subsection{Continuous Group Key Agreement\label{sec:CGKA}}
In the two-party case, a general notion of Continuous Key Agreement has been used to provide robust security guarantees such as forward secrecy and post-compromise security~\cite{perrin2013axolotl,perrin2016double,alwen2019double}.
For \Abrev{SGM}, this same notion has been extended as \Abrev{CGKA}, which is used as an abstraction for reasoning about the \Abrev{MLS} security guarantees~\cite{alwen2020security}.
A continuous group key-agreement scheme \CGKAdef\ is an abstract collection of algorithms used to maintain secure communications between a group of two or more participants, each algorithm being known and performed locally by participants:\\[2mm]
\centerline{$\CGKAdef = \left\{\,\Protocol{init},\, \Protocol{create},\, \Protocol{add},\, \Protocol{rem},\, \Protocol{upd},\, \Protocol{proc}\,\right\}$}\\[-3mm]

@@

These oracles define a security game \cite{shoup2004sequences} which emulates the \Abrev{ITM}.
Such security games are a common analysis tool made by cryptographers, where the ``game moves'' representing protocol semantics, and the objective is to show that the probability of an adversary \Adversary  breaking a given security property by playing this game is negligible.
Within the \CGKAsec, an adversary \Adversary is parameterized by \((T, C, N)\).
The \(T\) value indicates that the protocol runs in at most \(T\) epochs, during which time \(\mathcal{A}\) can make at most \(C\) ``challenge queries'' to the \Oracle{reveal}{} or \Oracle{chall}{} oracles, and the total unique group members of the protocol can never exceed \(N\).
The goal of \Adversary is to infer information about the shared TreeKEM root key.
Within the security game, this is modeled by \Adversary  ending the game via a challenge query to the \Oracle{chall}{} oracle, 
and subsequently demonstrating non-negligible advantage.
We model the TreeKEM protocol in terms of the \CGKAsec as the initial abstraction.

Advantage is formally defined by~\cite{alwen2020security}, but a brief definition goes as follows.
A random bit $b$, hidden from \Adversary, is flipped uniformly at random.
If $b = 1$ then the real TreeKEM protocol is used in the security game, otherwise the security game simulates TreeKEM with random data.
 \Adversary demonstrates advantage if and only if they can guess the value of $b$ with probability $\varepsilon > \frac{1}{2}$.
Since \Adversary can distiguish between random data and encrypted data, the princile of indistinguishability is violated and the protocol is not secure.

%[SSB] Its not clear that following subsections are abstractions. Either/both add a sentence here saying they are, or change subsection names to "Abstraction: ...". Maybe also mention here that 'this model needs blah hours to verify even for N=4' if you have such numbers.

@@

\subsection{Abstraction: Adversarial Direction}
While both the adversary \Adversary and each group member can initiate a new epoch within the \CGKAsec, modeling this explicitly unnecessarily complicates the model.
Recall that within the \Abrev{ITM}, \Adversary controls the network.
This is modeled within the \CGKAsec by messages sent from any source only being delivered if \Adversary permits it by querying \Oracle{deliver}{}.
Because the verification methodology will explore all possible interleaving of concurrent actions within the \CGKAsec, simulating nondeterministic computation, this can be utilized to form an additional simplifying abstraction.

Suppose for a given epoch $e_i$ that the non-empty set $M$ consists of all ``epoch advancing messages'' originating from calls to algorithms \Protocol{add}, \Protocol{rem}, and \Protocol{upd} by group members.
By the rules of the \CGKAsec, \Adversary selects some message $m \in M$ sent by group member $j$ and queries \Oracle{deliver}{g_j,\,g_k} to propagate delivery of $m$ to the group members and advance to the next communication epoch.
Without loss of generality, rather than group member $j$ creating and sending $m$ by calling either algorithm \Protocol{add}, \Protocol{rem}, or \Protocol{upd}, the adversary \Adversary equivalently queries the corresponding \Oracle{add-user}{g_j,\,g_k}, \Oracle{remove-user}{g_j,\,g_k}, or \Oracle{send-update}{g_j,\,g_k} oracle, respectively.
This operational equivalence permits the modeling of the \CGKAsec to only consider \Adversary--originating calls to algorithms \Protocol{add}, \Protocol{rem}, and \Protocol{upd} via queries to the oracles \Oracle{add-user}{}, \Oracle{remove-user}{}, and \Oracle{send-update}{}, respectively.
Consequently, the model does not need to consider any group member calls to \Protocol{add}, \Protocol{rem}, and \Protocol{upd}, allowing these states to be safely removed from the model encoding without effecting protocol semantics.
The model can explicitly track epoch progress by the adversary via $t$ and $c$ variables corresponding to epoch index and a binary indicator of a challenge query made within the current epoch, respectively.
%The application of this simplifying abstraction composed with the original \CGKAsec is depicted in Figure \ref{fig:CGKA-Adversarial-Driven}.
% [SS] I think the main point of above paragraph is that modeling/proving by epoch avoids the need for non-A originating calls? But I'm not sure its clear to reader. 
%[AW] Some more descriptive language here.
%[SSB] This sounds like partial order reduction (enabled by default in SPIN), but can be disabled using -DNOREDUCE. Presumably our solution improved over SPIN - assuming so add sentence here saying so. Also check SPIN docs to ensure that its actually using the reduction (it doesnt in some cases).

@@

\subsection{Abstraction: Parameterized Bounds of $T$, $C$, and $N$}

@@

\subsection{Abstraction: Required Progression\label{sec:abstraction-progression}}

@@

\subsection{Abstraction: Markovian Property}

@@

The \Promela encoding \footnote{Source code: \url{https://github.com/recursion-ninja/Manuscript-2024-SBMF}} of \CGKAmodN{N} produces all possible states and transitions of TreeKEM, and this representation is then supplied to a model checker for verification.

The explicit state model checker Spin~\cite{HolzmannSpinBook2003} was chosen as the verification tool for this work.
As an explicit-state model checker, Spin utilizes a representation of all possible model states as a \Abrev{FSM}~\cite{clarke1981design}, with transitions corresponding to possible changes in model state.
Spin then verifies a user-specified property by traversing through all transitions of the \Abrev{FSM}, and checking that that a property holds at each model state.
%~\cite{Vardi1986}.
The principle drawback of explicit state model checking is the immense state-space resulting as the model complexity grows~\cite{burch1992symbolic}.
As a venerable tool, Spin does much to address this~\cite{rudin1987limits} but the authors' own encoding adaptations which further mitigate the state-space explosion are detailed below in Sections \ref{sec:bit-array} and \ref{sec:popcount}.
Additionally, Spin supports numerous compile time directive which alter the state-space representation, search trajectories, and time-memory trade-offs (see Table \ref{tab:spin-confg}).

@@

In the default \Promela type system, an array of bits is encoded as an array of bytes at runtime, resulting in each ``bit value'' requiring 8 bits during verification.
This encoding inflates the state-vector size unnecessarily by 7 bits additional bits and causes the model checker to consider 256 states instead of the actual 2 binary states which are valid encodings ($\texttt{0x00} = \textsc{True}$, $\texttt{0xFF} = \textsc{False}$).
Consequently, verification times are slower, and at least eight times more memory is required.
To address this, 
bits within a fixed bit-width \Promela type are manually toggled using C Pre-processor macros,
% [SS] Check that my rephrasing is correct
% [AW] Yes!
%the manual toggling bits within a fixed bit-width \Promela type using C Pre-processor (CPP) macros is included, 
reducing the state-vector size and improving runtime efficiency while minimizing memory usage.

@@

The authors included PopCount as a small, inlined sequence of arithmetic and bit-wise operations which runs in \BigO{1} time with no additional memory.
Substituting PopCount in place of the naive iteration has two notable effects, reducing the state-vector length by removing the loop counter and reducing the number of state transitions by removing the entire control flow structure of the loop.
For a bit-array of length $n$, the number of transitions encoded in the model by iteration increases exponentially with respect to $n$. This exponential transition encoding occurs because for each iteration of the bit counting loop, the set bit counter's value can either be incremented or maintained, hence $2^{n}$ transitions to count the number of set bits.
Conversely, the PopCount macro derives a \emph{constant} number of transitions independent of $n$.

@@

%[SSB] what does UnderNoteRef mean? If there is a point to it, define what it means; else get rid of it. 
\begin{definition}[Forward Secrecy with Updates]
If the state of any group member is leaked during an epoch\UnderNoteRef{1}, the shared keys from all previous epochs remain hidden from the adversary\UnderNoteRef{2}.
\end{definition}%
\noindent\textbf{LTL Predicate} \LTLPredicate{FSU}\textbf{:}%
{\small%
\[%
\Box \big(\, \underset{\UnderNote{1}}{\texttt{CGKA@move\_corrupt}} \implies \underset{\UnderNote{2}}{\texttt{learnedLegacyKey}} = \bot \,\big)%
\]%
}

The \Promela model is named \TermLTL{CGKA}, and the state within the model at which \Adversary queries the oracle  \Oracle{corr}{\TermLTL{targetID}} is labeled \TermLTL{move\_corrupt}.
The term \TermLTL{CGKA@move\_corrupt} will be true only when \TermLTL{CGKA} is in state \TermLTL{move\_corrupt}.
Additionally, the model contains a boolean flag \TermLTL{learnedLegacyKey} indicating when \Adversary has learned, during the \emph{current} epoch, information about the shared group key of any \emph{any previous} epoch.
The \Abrev{LTL} formula \LTLPredicate{FSU} is a direct translation of the natural language \Abrev{FSU} description.

@@

The methodology presented is two-fold.
First, the performance is measured of the model \CGKAmodN{N} described in this work and compared with the performance of a precursor model defined by~\cite{washburn2022formal} 
denoted as \CGKAmodNT{N}{T}.
%[SSB] after "described" add "in sections 2.2-2.6"
The comparison model \CGKAmodNT{N}{T} is the same semantic encoding of \CGKAmodN{N} \emph{before} the abstractions described were applied, i.e:  $$\Code{Abstractions}\Parens{\CGKAmodNT{N}{T}} =  \CGKAmodN{N}$$
Therefore \CGKAmodNT{N}{T} is an apt choice as a comparative baseline from which the effects of the composed abstractions can be measured.
Comparison of performance is made across several metrics, including state-vector length, runtime, memory usage, state-space size, and transition count.
Both models performed verification of the \LTLPredicate{FSU} and \LTLPredicate{PCS} properties.
Second, \CGKAmodN{N} is utilized to verify \LTLPredicate{FSU} and \LTLPredicate{PCS} for parameters far beyond the capabilities of \CGKAmodNT{N}{T}.
Subsequently, the authors provide brief contextual commentary accompanying these verification results.

@@

%\centering

@@

@@

& \multicolumn{2}{c}{\TableHeadFontSize\parbox[][6mm][c]{14mm}{ \CGKAmodN{N}}}
& \multicolumn{2}{c}{\TableHeadFontSize\parbox[][6mm][c]{19mm}{\centering \CGKAmodNT{N}{T}}}\\

@@

& \parbox[t]{5mm}{\raggedleft$T\:$}
& \parbox[t]{9mm}{\scriptsize \centering State\\Vector}
& \parbox[t]{6mm}{\raggedleft$T$}
& \parbox[t]{12mm}{\scriptsize\centering State\\Vector}\\

@@

\hspace*{9mm}

@@

& \parbox[t]{10mm}{\TableHeadFontSize\centering $N$}

@@

Given the new range of tractable verification provided by \CGKAmodN{N}, verification was performed with model parameters which were previously untenable using \CGKAmodNT{N}{T}. 
Verification of  \Abrev{FSU} and \Abrev{PCS} was conducted on the models \CGKAmodN{16}.
Both models had a state-vector of $100$\siBytes.
Then verification results produced by  \CGKAmodN{16} for both properties were \emph{positive}, affirming that the properties hold for the model.
The positive verification of the LTL encodings for \Abrev{FSU} and \Abrev{PCS} given by \CGKAmodN{16} are not surprising.
Rather, it is consistent with the expectations shown by the previous \CGKAdef abstraction formalization work~\cite{alwen2020security}.
While consistent, this is the first instance known to the authors which has modeled the TreeKEM protocol and verified \Abrev{FSU} and \Abrev{PCS} for all $(\,T, C, N\,)$--Adversaries, with $T \in \NumericRangeOpenR{1}{\infty}$, $C \in \NumericRange{1}{T}$, and $N \in \NumericRange{2}{16}$.
Formal verification for TreeKEM of this scope has not been done before.

@@

This work presents a verification of \Abrev{FSU} and \Abrev{PCS} for TreeKEM; an important undertaking in establishing the security of and confidence in the protocol.
Furthermore, an example of effective, and generally applicable model abstraction and model encoding techniques is illustrated.
The authors believe that, though no abstraction in isolation is novel, this is the first work to compose these abstractions and encoding techniques in the combination presented.
The exploration of combining these techniques has been shown necessary for achieving scalability when verifying a complex model such as the TreeKEM protocol.
An unqualified improvement was obtained across each of the following measurable dimensions: model complexity,  parameterization, state-vector length, and verification space as well as time.
These improvements were essential to perform verification of the largest known TreeKEM group size to date.
While the application and analysis of presented techniques herein was limited to TreeKEM, the techniques are not so bespoke that they apply only to modeling and verifying TreeKEM.
Rather, it is probable that only minor creativity is require to transfer them to other modeling efforts.
